---
title: "Integrated Exercise"
output: html_document
---

<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.6.0/css/all.css" integrity="sha384-aOkxzJ5uQz7WBObEZcHvV5JvRW3TUc2rNPA7pe3AwnsUohiw1Vj2Rgx2KSOkF5+h" crossorigin="anonymous">

# Introduction

During this workshop you have learned to import, manipulate, visualise
and analyse data. Time to use all these skills and create a research
report!

This research report will focus upon the relation of motivation with
math scores. We will use data that was collected in 2010 during a large
project on the transition of secondary to higher education or the labour
market. We followed almost 4000 students for two years and collected
data during 5 waves. If you want to know more about the project, you can
have a look at
[this](https://www.sciencedirect.com/science/article/pii/S104160801500059X?casa_token=XAxx5GZKHCsAAAAA:B2Xt82w-rWRH6q2KocvVOTOJxIX3hxSGLfwnKZyh8_xUbUFDl7_HQZdvR8zwLTYuqElddNq-0rc)
or
[this](https://journals.sfu.ca/flr/index.php/journal/article/view/647)
article. Today, we only use the data collected during the first two
measurement waves. Students were in their final year of secondary
education.

The data set contains information on 15 variables:\
- *ID_School*: ID of the school\
- *ID_Class*: ID of the class\
- *ID_Student*: ID of the student\
- *Educ_level*: educational track of the studion [(1: general education,
2: technical education, 3: vocational education, 4: arts
education)]{style="font-size:10;"}\
- *Gender*: gender of student (0: female, 1: male)\
- *SES_language*: language indicator of social-economic status that was
used by the government to indicate whether or not a student was at risk
(0: does not apply, 1: applies)\
- *SES_neighbourhood*: neighbourhood indicator of social-economic status
that was used by the government to indicate whether or not a student was
at risk (0: does not apply, 1: applies)\
- *SES_educ_level*: education indicator of social-economic status that
was used by the government to indicate whether or not a student was at
risk (0: does not apply, 1: applies)\
- *IQ_score*: raw score on a non-verbal IQ-test (min. score: 0, max.
score: 40) - *Math_score*: raw score on a math test (problem-solving)
(min. score: 0, max. score: 24)\
- *Reading_score*: raw score on a reading comprehension test (min.
score: 0, max. score: 24)\
- *Autonomous_motivation*: autonomous motivation for studying (based on
average score on 6 items, min. score: 0, max. score: 5)\
- *Controlled_motivation*: controlled motivation for studying (based on
average score on 6 items, min. score: 0, max. score: 5)\
- *Amotivation*: amotivation for studying (based on average score on 3
items, min. score: 0, max. score: 5)

💡 *Don't forget to load the necessary packages: `tidyverse` and
`here`.*

```{r}
#| echo: true

# CODE HERE

```

# 1. Import and clean data

## 1.1 Import the data

Import the data

```{r}
#| echo: true

# CODE HERE

```

## 1.2 Check data

Check all variables:

-   Are they correctly defined (numeric, character, ...)?
-   Which variables should be converted to another 'type' (numeric,
    character, ...)?

```{r}
#| echo: true

# CODE HERE

```

*What is your conclusion?*

Most categorical variables (`ID_Class`, `ID_Student`, `Educ_level`,
`Gender`, `SES_language`, `SES_neighbourhood`, `SES_educ_level`) are
defined as 'integer'. These variables should be converted to type
'character'. The variable `X` contains row numbers and should be
removed.

You might have already noticed that several variables contain missing
values (NA's). However, it is always save to check this for all
variables and also to look for 'strange' values. **Get an overview of
the variables using the function `summary()`.**

```{r}
#| echo: true

# CODE HERE

```

*What is your conclusion?*

Almost all variables have missing values. The maximum value of the
motivation variables (`Autonomous_motivation`, `Controlled_motivation`,
`Amotivation`) is 99. This is, of course, an impossible value (maximum
was 5). To be sure that these variables don't contain other strange
values, **you make a quick barplot of the distribution of each of these
variables. You save the plots as objects `P1`, `P2` and `P3`.**\
(A barplot is most of the time not a good choice to visualise
quantitative variables. However, to get an idea of every value in a
quantitative variable, a barplot is more informative than a histogram or
density plot.)

💡 *Attention! You have to create three barplots. These plots can be
combined into one plot using the package `patchwork`. You have to load
this package and add code to combine the three plots. This code has
already been included in the code block below. This code now has a `#`
in front of it. Remove that `#` so the code will be executed.*

```{r}
#| echo: true

# CODE HERE

```

*What is your conclusion?*

The motivation variables (`Autonomous_motivation`,
`Controlled_motivation`, `Amotivation`) don't contain other strange
values.

**You decide to also visualise the distribution of the categorical
variables: `Educ_level`, `Gender`, `SES_language`, `SES_neighbourhood`
en `SES_educ_level` (using a barplot).**

You already know that these variables are included as numeric variables,
while they are of type character. To make sure that you get correct
values on the x-axis of your barplots, specify the `x aesthetic` as
follows `ggplot(aes(x = as.character(Educ_level)))`.

💡 *Attention! You have to create five barplots. These plots can be
combined into one plot using the package `patchwork`. You have to add
code to combine the five plots. This code has already been included in
the code block below. This code now has a `#` in front of it. Remove
that `#` so the code will be executed.*

```{r}
#| echo: true

# CODE HERE

```

*What is your conclusion?*

The distribution of the five categorical variables looks fine. (They all
have missing values!) Category 4 of the variable `Educ_level` contains
little observations compared to that of the three other categories. It
would be nice if the categories of all these variables would be
appropriately labelled. (In that case, you would immediately know what
category 4 of the variable `Educ_level` refers to.)

## 1.3 Clean data

After having checked the data, you know that the following actions needs
to be performed:

1.  convert all categorical variables to a character variable
2.  label the categories of the variables `Educ_level`, `Gender`,
    `SES_language`, `SES_neighbourhood` and `SES_educ_level` and store
    them as new variables
3.  recode the values 99 in the motivation variables
    (`Autonomous_motivation`, `Controlled_motivation`, and
    `Amotivation`) to NA and store them as new variables
4.  remove the first column

Write the code to perform the actions listed above and save as a new
data frame `transition_data_cleaned`

```{r}
#| echo: true

# CODE HERE

```

**Don't forget to check your recoding work!**

First, *check whether each variable is defined correctly. (Are all
categorical variables defined as a character variable? Is the redundant
variable removed from the data?)*

```{r}
#| echo: true

# CODE HERE

```

*Check whether you labelled the categorical variables correct using the
function `table()`.*

```{r}
#| echo: true

# CODE HERE

```

## 1.4 Prepare data for analysis

Create a data frame `transition_data_for_analysis` that only contains
the following variables:

-   ID_School\
-   ID_Class\
-   ID_Student\
-   Educ_level and Educ_level_cat\
-   Gender and Gender_cat\
-   all SES-indicators (original and recoded variables)\
-   IQ_score, Reading_score, Wisk_score\
-   the motivation variables (only the ones that you recoded)

To select these variables efficiently, the 'selection helpers' of the
select()-function are handy. More information on these helpers can be
found [here](https://dplyr.tidyverse.org/reference/select.html). Also
include a standardized version of every quantitative variable in your
data set. To standardize a variable, you can use the function `scale()`.
Finally, you decide to exclude all students in the arts education track
(based on the results of the barplot you created in 1.2).

```{r}
#| echo: true

# CODE HERE

```

Check you new data frame `transition_data_for_analysis`.

```{r}
#| echo: true

# CODE HERE

```

Check if the data is filtered correctly.

```{r}
#| echo: true

# CODE HERE

```

# 3. Some descriptives

Before doing the analysis, you want to check some things.

First, **you want to get an idea of the number of observations per
school.** To do so, you create a lollipop plot. To help the reader of
your rapport, you make sure that the 'bars' are arranged in ascending
order. Color bars with counts less than 20 in red. (You first have to
create a new variable `color_ID` to be avble to do that!) Use your
ggplot2-skills to make the plot "publication-ready".

```{r}
#| echo: true

# CODE HERE

```

You also create a table with the descriptive statistics (mean, SD, min,
max) of the raw (unstandardized) variables `IQ_score`, `Reading_score`
and `Math_score`. To so, you first create a vector per variable (using
the function `c()`) that contains those four values. To calculate these
values you use the functions `mean()`, `sd()`, `min()` and `max()`.
Then, you combine the three vectors into one tibble (`descriptives`)
using the function `bind_rows()`. (This function pastes the rows below
each other.) Don't forget to print the resulting tibble.

```{r}

# CODE HERE

```

Before turning the tibble into a real table, **you must round all
numbers to two places after the decimal** combining the functions
`mutate_all` and `round()`. After that, you also **create a new column
`Name` that contains the names of the variables. Make sure it appears
before the column `mean`.**

```{r}

# CODE HERE

```

Reuse the code you wrote and turn your tibble into a real table by using
the function `kable()`.\*\* Check out the help-page of the
`kable()`-function to find out how you can align the five columns!
Center all but the first columns. (The first one can be "left" aligned.)

💡 *Attention! To use the function `kable()` you have to install and
load the package `knitr`.*

```{r}
#| echo: true 

# CODE HERE

```

Much more options are available when you also install the package
`kableExtra`. This website provides more information on the use of that
package:
https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html
One of the many nice features of kableExtra are the built-in themes:\
- kable_classic()\
- kable_classic2()\
- kable_paper()\
- kable_minimal()\
- kable_material()\
- kable_material_dark()

Reuse the code your wrote above and add a theme. Play around with these
themes to get to know them.

💡 *Attention! To use these functions you have to install and load the
package `kableExtra`.*

```{r}
#| echo: true

# CODE HERE

```

# 4. Analyse the data

**Fit a first linear regression model with math score as dependent
variable and the thee motivation variables as independent variables.**

Use standardized variables. Save your model as an object `Model_lm1` and
print the model summary.

```{r}
#| echo: true
 
# CODE HERE

```

**Fit a second linear regression model that builds on the previous one
but adds gender, educational level, IQ-score and reading-score as
control variables.**

Use standardized variables. Save your model as an object `Model_lm2` and
print the model summary.

```{r}
#| echo: true

# CODE HERE

```

Before reporting on the results, \*\*you have to check the assumptions
of both models. Several R-packages perform these kind of tasks. One
example is the `performance` package. Install this package and use the
function `check_model()` to do a visual check of the assumptions.

💡 *Attention! To use the function `check_model()` you have to install
and load the package `performance`.*

Check the assumptions of the model `Model_lm1` first.

```{r}
#| echo: true

# CODE HERE

```

Do the same for the other model.

```{r}
#| echo: true

# CODE HERE

```

These results look quite good. However, you also want to do an
additional check for outliers. Go to [this
website](https://easystats.github.io/performance/index.html) and look
for the function that you need to do an additional check on outliers.

```{r}
#| echo: true

# CODE HERE

```

Another important assumption concerns the distribution of the dependent
variable (should be normally distributed). You do not need a package to
check on that distribution, but **use your ggplot2-skills to make a
histogram of the distribution of the dependent variable.** 'Style' the
histogram so you can use it in a publication.

```{r}

# CODE HERE

```

The tail on the left-hand side of the distribution worries you a bit.
You decide to discuss this with your supervisor.

To finish up the report, you **create a nice table of the estimates of
both models**. To create this table, you use the function `tab_model()`
of the package `sjPlot`. Go the help-page of this function and figure
out how to turn off the printing of p-values, print a 90% confidence
interval, rename the column names to 'Variables', 'Est.' and '90% CI'
and remove the brackets around Intercept.

💡 *Attention! To use the function `tab_model()` you have to install and
load the package `sjPlot`.*

```{r}

# CODE HERE
 
```

Before sending the report to your supervisor, you compare both models.
To do so, you first **make a data set `transition_data_omitted` that
excludes all missing values for the variables included in `Model_lm2`.**
(This is necessary in order to compare the models.)

```{r}

# CODE HERE

```

Then, you **refit both models using the new data set.**

```{r}

# CODE HERE

```

Finally, you **compare both models using the function
`compare_performance`** of the package `performance`.

```{r}

# CODE HERE

```

To be able to sent the report to your supervisor, you render it as an
html-file

::: callout-tip
## A Quarto tip!

Please note, if you Render the document to send it to someone else you
want to have all the information in one single file so that you can send
the .html file to that person. To make sure all the figures etc are in
the file, specify the option `self-contained: true` in the YAML. See it
in action in the .qmd file for this exercise.
:::
